Type Ia Supernova Explosions from Hybrid Carbon-Oxygen-Neon White Dwarf Progenitors That Have Mixed During Cooling

### 2019-08-31 paper/reply_to_referee.txt

We thank the referee for reading out manuscript and offering very
constructive criticism and suggestion. We are in almost complete agreement
with the referee and have modified the manuscript accordingly. We address
our changes for each point below. As a result of our revisions, we believe 
the manuscript is stronger and we hope it is now ready for publication.

> *Methodology*

> - It is not clear to me if the input deflagration speed (pg 7) has been
> recalibrated to the particular initial composition of the mixed hybrid
> core. Willcox et al (2016) just reports recomputing detonations, but there
> is no mention of deflagrations (at least I have not been able to find it).

We regret no explaining the methodology of our deflagration speeds. The 
answer is that the deflagration speeds for input to the flame model are 
from tabulated results. Laminar flame speeds are first 
obtained by interpolation within a
three-dimensional table of 12C, 22Ne, and log density obtained from
prior direct numerical simulations. Here 22Ne is uses as a proxy for
for neutron-rich elements so the interpolated laminar flame speeds capture
the boost due to neutronization. These laminar flame speed are
then boosted to account for the effects of unresolved buoyancy
and background turbulence and passed to the flame model. We have
updated the text of the manuscript to explain the flame speeds
and included a citation to earlier work describing the tabulated
flame speeds in detail.
[NOTE: CHECK THIS. The models include 20Ne. Does the above
description properly account for that? As I recall, nothing new
was done to the tables for the hybrid Willcox et al. models.]

Also, the detonation tests in Willcox et al. were tests of the
energy release of the flame model and not flame speed tests. The 
model identical amounts of energy for both deflagrations and 
detonations in identical material, so detonation simulations were
a simple check of the modifications we made to the model to
include 20Ne in the mix.

@@

> - Why is the MESA calculation stopped and mapped to FLASH just when
> the central temperature reaches log Tc = 8.9? What is the criterion for
> choosing such a temperature? Does it depend on composition?

Again, we did not explain our methodology well. This temperature is
very close to that required to ignite the deflagration of the
first phase of the explosion, so the progenitor is in a ready 
for ignition state. Yes, it is a (somewhat mildly) a function of 
composition, but at this temperature these models have very little 
evolution left before the flame would be born even with differing 
composition. We have updated the text to better explain this issue.

@@

> - Why was the C+O model constructed with the same central temperature and
> density as the mixed hybrid model? Would it be more appropriate to follow
> a C+O white dwarf with MESA until some runaway condition was accomplished?

> - Related to the last point above: it is not clear how was assigned the
> composition profile to the C+O model. In particular, why is the abundance
> of 22Ne in the C+O model smaller than that in the hybrid mixed model? Does
> it has an impact on the final yield of 56Ni?

The referee raises an interesting point about how one might perform
these parameter studies. Indeed, a more realistic C/O progenitor as
would be produced by MESA would be "state of the art." The parameterized
C/O models we chose are motivated by results from stellar evolution and,
in particular, have compositions that reflect the effect of ``simmering", 
that is slow convective burning, in the late stages of evolution as the
WD approaches the Chandrasekhar limit. This effect is observed as
the lower C fraction interior to about r =1.2e8 cm (the solid curve
in the bottom panel of Figure 1). Accordingly, while parameterized, 
the structure of C/O progenitor does reflect our current understanding 
of the state just prior to ignition.

For this study we decided to use these parameterized C-O models because
being able to construct them to match the central density and 
temperature of the hybrid models allowed us to better control the 
differences between the hybrid and traditional models. More realistic
C-O models would not match the conditions of the hybrid models, and
as we have shown in previous work, differences in density can have
a substantial effect on the yields.

We have added text to the manuscript and slightly re-arranged the 
description to better explain the motivation for our choice.

### 2019-08-31 paper/ms.tex
More work on the new hybrid model manuscript to address the referee's concerns.

The reaction progress variable is $\phi$ and it varies from $\phi=0$ for
unburned fuel to $\phi=1$ for burned ash. $\phi$ is evolved
via an advection-diffusion-reaction equation,
\begin{equation}
  \label{eq:ard}
  \partial_t \pv + \vec{u}\cdot\nabla \pv = \kappa \nabla^2 \pv +
\frac{1}{\tau} R\left(\phi\right) ,
\end{equation}
where $\vec{u}$ is the velocity of the fluid, $\kappa$ is the
diffusion coefficient, $\tau$ is the reaction timescale, and $R(\phi)$ is
a non-dimensional function describing the reaction. The parameters
$\kappa$, $\tau$, and $R(\phi)$ are tuned to propagate the reaction
front at a prescribed speed.  Our model uses the ``sharpened KPP''
 described by~\cite{VladWeirRyzh06},
with $R\propto(\phi-\epsilon)(1-\phi+\epsilon)$, where
$\epsilon \simeq 10^{-3}$.  This scheme has been shown to be
acoustically quiet, stable, and to give a unique flame speed
\citep{townsley.calder.ea:flame}. The input flame speeds come from
tabulated results obtained by direct numerical simulations of
thermonuclear burning.
{\color{blue}
The flame speeds are obtained via linear interpolation within a 
three-dimensional table that combines the results of \citet{timmes92} 
and \citet{Chametal08}, and the dimensions of the table are \C{12} 
mass fraction, \Ne{22} mass fraction and log of density.}
These laminar flame speeds are boosted to account for the speed-up 
of burning due to unresolved buoyancy and background turbulence~\citep{Khok95,
gamezo.khokhlov.ea:thermonuclear,townsley.calder.ea:flame,jacketal2014}.

@@

The hybrid white dwarf model that served as the initial condition
for the simulations of supernova explosions presented in this work
was constructed with the one-dimensional stellar evolution code
MESA~\citep{mesa1,mesa2,mesa3,mesa3e}. The evolution of the model 
included larger nuclear reaction networks than previous studies and 
thus resulted in a electron-to-baryon gradient that became unstable 
to mixing as the interior cooled, leading to a lower central C 
fraction than previous hybrid models~\citep{brooksetal2017}. We 
selected the 1.09 \Msun WD model 
with the 0.4 \Msun C/O core from \citep{brooksetal2017}.
In order to grow this WD towards the Chandrasekhar mass, it was inserted into 
a binary with a 1.4 \Msun He star donor having a 3 hour orbital period, similar to 
the simulations in \citet{brooksetal2016},
and the system was
allowed to evolve.  The WD grows, experiences central carbon ignition and 
simmering, and then is stopped when its central temperature reaches 
$\log T_{\rm c}(\Kelvin) = 8.9$, {\color{blue} at which point the 
temperature is high enough to ignite the deflagration that is the first
stage of the explosion.}

@@

The simulations performed for this study consisted of a suite of
30 two-dimensional simulations of
thermonuclear supernova
explosions from hybrid C-O-Ne progenitors in the DDT explosion paradigm. These
were compared to a suite of supernova simulations from traditional
C-O progenitors. In both suites, the simulation
begins with a progenitor model mapped to the two-dimensional
\FLASH\ grid. The burning is initiated with a ``match head,'' a region
in the white dwarf's center that is fully burned to nuclear statistical
equilibrium (NSE).
This initially burned region ignites a deflagration, a subsonic
flame, and because the match head was perturbed it is unstable to
the Rayleigh-Taylor instability with the result that buoyant plumes
rise. The star is partially consumed during this deflagration phase,
and the star responds by expanding.
When a plume reaches a specified density,
a detonation is initiated, and the simulation continues until
the expanding star reaches low densities at which point burning
effectively ends. This section provide details
of the implementation of this method.

@@

>>>>

The reaction progress variable is $\phi$ and it varies from $\phi=0$ for
unburned fuel to $\phi=1$ for burned ash. $\phi$ is evolved
via an advection-diffusion-reaction equation,
\begin{equation}
  \label{eq:ard}
  \partial_t \pv + \vec{u}\cdot\nabla \pv = \kappa \nabla^2 \pv +
\frac{1}{\tau} R\left(\phi\right) ,
\end{equation}
where $\vec{u}$ is the velocity of the fluid, $\kappa$ is the
diffusion coefficient, $\tau$ is the reaction timescale, and $R(\phi)$ is
a non-dimensional function describing the reaction. The parameters
$\kappa$, $\tau$, and $R(\phi)$ are tuned to propagate the reaction
front at a prescribed speed.  Our model uses the ``sharpened KPP''
 described by~\cite{VladWeirRyzh06},
with $R\propto(\phi-\epsilon)(1-\phi+\epsilon)$, where
$\epsilon \simeq 10^{-3}$.  This scheme has been shown to be
acoustically quiet, stable, and to give a unique flame speed
\citep{townsley.calder.ea:flame}. The input flame speeds come from
tabulated results obtained by direct numerical simulations of
thermonuclear burning.
{\color{blue}
The flame speeds are obtained via linear interpolation within a 
three-dimensional table that combines the results of \citet{timmes92} 
and \citet{Chametal08}, and the dimensions of the table are \C{12} 
mass fraction, \Ne{22} mass fraction and log of density. While the
WD contains additional species, we have shown that treating the
abundance of \Ne{22} as a proxy for neutron-rich elements captures the
speed-up of laminar flames due to neutronization and thus reasonably 
produces variations in flame speed due to composition~\cite{jacketal2010}.}
These laminar flame speeds are boosted to account for the speed-up 
of burning due to unresolved buoyancy and background turbulence~\citep{Khok95,
gamezo.khokhlov.ea:thermonuclear,townsley.calder.ea:flame,jacketal2014}.

@@

The hybrid white dwarf model that served as the initial condition
for the simulations of supernova explosions presented in this work
was constructed with the one-dimensional stellar evolution code
MESA~\citep{mesa1,mesa2,mesa3,mesa3e}. The evolution of the model 
included larger nuclear reaction networks than previous studies and 
thus resulted in a electron-to-baryon gradient that became unstable 
to mixing as the interior cooled, leading to a lower central C 
fraction than previous hybrid models~\citep{brooksetal2017}. We 
selected the 1.09 \Msun WD model 
with the 0.4 \Msun C/O core from \citep{brooksetal2017}.
In order to grow this WD towards the Chandrasekhar mass, it was inserted into 
a binary with a 1.4 \Msun He star donor having a 3 hour orbital period, similar to 
the simulations in \citet{brooksetal2016},
and the system was
allowed to evolve.  The WD grows, experiences central carbon ignition and 
simmering, and then is stopped when its central temperature reaches 
$\log T_{\rm c}(\Kelvin) = 8.9$, {\color{blue} at which point the 
temperature is just about high enough to ignite the deflagration that is 
the first stage of the explosion.}

@@

The simulations performed for this study consisted of a suite of
30 two-dimensional simulations of thermonuclear supernova explosions 
from hybrid C-O-Ne progenitors in the DDT explosion paradigm. These
were compared to a suite of supernova simulations from traditional C-O 
progenitors. {\color{blue} The C-O progenitors are parameterized and 
include the effects of convective ``simmering" in the core as the WD 
approaches the Chandrasekhar mass~\citep{Chametal08,PiroBild08,jacketal2010}.
Using parameterized C-O models allowed us to choose conditions, e.g.
central density, to control differences between the hybrid and traditional 
models and thus assess the effect of the hybrid structure.
}

@@

In both suites, the simulation
begins with a progenitor model mapped to the two-dimensional
\FLASH\ grid. The burning is initiated with a ``match head,'' a region
in the white dwarf's center that is fully burned to nuclear statistical
equilibrium (NSE).
This initially burned region ignites a deflagration, a subsonic
flame, and because the match head was perturbed it is unstable to
the Rayleigh-Taylor instability with the result that buoyant plumes
rise. The star is partially consumed during this deflagration phase,
and the star responds by expanding.
When a plume reaches a specified density,
a detonation is initiated, and the simulation continues until
the expanding star reaches low densities at which point burning
effectively ends. This section provide details
of the implementation of this method.
